---
title: "Regression and Regression Tree Tutorial"
author: "Xiao Liu"
date: "Jan. 30, 2017"
output: 
  html_document:
    number_sections: yes
    toc: yes
    fig_width: 15
    fig_height: 10
---
# Setup and Import Insurance dataset  
```{r Setup and data import}

# Load the following packages. Install them first if necessary.

library(caret)
library(psych)
library(rpart)
library(rpart.plot)
library(RWeka)
library(rminer)
library(matrixStats)
library(knitr)

# import data
setwd("~/Desktop")
insurance <- read.csv(file = "insurance.csv", stringsAsFactors = TRUE)
# Overall inspection
str(insurance)
summary(insurance)
```

# Data exploration: some examples

```{r Data exploration}
# histogram of insurance expenses
hist(insurance$expenses)

# correlations

# exploring relationships among features: correlation matrix
cor(insurance[c("age", "bmi", "children", "expenses")])
# visualizing correlations
pairs.panels(insurance)

```

# Building a base explanatory model on the whole data set

```{r A base explanatory model}
ins_base_model <- lm(expenses ~ age + children + bmi + sex + smoker + region, data = insurance)
# this is equivalent
ins_base_model <- lm(insurance[,7] ~ ., data = insurance[,-7]) 

# exam the model and performance summary
ins_base_model
summary(ins_base_model)

```
# Build a base predictive model and evaluation prediction performance  
 
```{r Train a predictive model}
set.seed(500)
inTrain <- createDataPartition(y=insurance$expenses, p = 0.70, list=FALSE)
train_target <- insurance[inTrain,7]
test_target <- insurance[-inTrain,7]
train_input <- insurance[inTrain,-7]
test_input <- insurance[-inTrain,-7]

#Build a model using training dataset

ins_base_train_model <- lm(train_target~., data = train_input)
ins_base_train_model
# see more detail about the estimated beta coefficients
summary(ins_base_model)
summary(ins_base_train_model)

```
# predict on the test data based on the trained base model

```{r Apply predictive model}

predictions_base_test <- predict(ins_base_train_model, test_input)

# compare the correlation between acutal and predicted expenses in test data
summary(predictions_base_test)
summary(test_target)

# compare the correlation
cor(predictions_base_test, test_target)

# compare the correlation between actual and predicted expneses in training data
predictions_base_train <- predict(ins_base_train_model, train_input)
cor(predictions_base_train, train_target)
```

# Use rminer to generate model's evaluation metrics
```{r Generate performance metrics}
# Generating multiple prediction evaluation metrics using rminer package
# performance of predictions on testing data 
mmetric(test_target,predictions_base_test,c("MAE","RMSE","MAPE","RMSPE","RAE", "RRSE", "COR", "R2"))
# performance of predictions on training data
mmetric(train_target,predictions_base_train,c("MAE","RMSE","MAPE","RMSPE","RAE", "RRSE", "COR", "R2"))
```
# Regression Trees and Model Trees 
```{r Train an rpart model}
# regression tree using rpart
ins_rpart_model <- rpart(train_target ~ ., data = train_input)
?rpart
# get basic information about the tree
ins_rpart_model
# get more detailed information about the tree
summary(ins_rpart_model)
# use the rpart.plot package to create a visualization
# a basic decision tree diagram
rpart.plot(ins_rpart_model, digits = 3)
# a few adjustments to the diagram
#The fallen.leaves parameter forces the leaf nodes to be aligned at the bottom of the plot, while the type and extra parameters affect the way the decisions and nodes are labeled
# rpart.plot(ins_rpart_model, digits = 4, fallen.leaves = TRUE, type = 3, extra = 101)
```
# Evaluate rpart model
```{r Apply and evaluate an rpart model}
# generate predictions for the testing and training dataset
predictions_rpart_test <- predict(ins_rpart_model, test_input)
predictions_rpart_train <- predict(ins_rpart_model, train_input)
# compare the distribution of predicted values vs. actual values in testing data
summary(predictions_rpart_test)
summary(test_target)
# compare the correlation between actual and predicted expenses in testing data
cor(predictions_rpart_test,test_target)
# compare the distribution of predicted values vs. actual values in training data
summary(predictions_rpart_train)
summary(train_target)
# compare the correlation
cor(predictions_rpart_train,train_target)
```
# Generate prediction performance metrics using rminer package  
```{r Generate rminer metrics}
# Performance of predictions on test data
mmetric(test_target,predictions_rpart_test,c("MAE","RMSE","MAPE","RMSPE","RAE", "RRSE","COR", "R2"))
# Performance of predictions on train data
mmetric(train_target,predictions_rpart_train,c("MAE","RMSE","MAPE","RMSPE","RAE", "RRSE","COR", "R2"))
```

# Model tree - M5P (RWeka)

```{r Train an M5P Model Tree}
WPM("list-packages", "installed")
ins_m5p_model <- M5P(train_target ~ ., data = train_input)
# display the tree
ins_m5p_model
# generate the summary of the model
summary(ins_m5p_model)
```
# M5P : prediction and metric evaluation
```{r Apply and evaluate an M5P model}
# generate predictions for the model
predictions_m5p_test <- predict(ins_m5p_model, test_input)
# Generating prediction performance metrics using rminer package
# Performance of predictions on test data
mmetric(test_target,predictions_m5p_test,c("MAE","RMSE","MAPE","RMSPE","RAE","RRSE","COR","R2"))

```

# Improve lm model performance by interaction and quadratic terms

```{r Improve lm model}
# add a higher-order "age" term
insurance$age2 <- insurance$age^2
# add an indicator for BMI >= 30
insurance$bmi30 <- ifelse(insurance$bmi >= 30, 1, 0)
#Partition the dataset again
set.seed(500)
inTrain <- createDataPartition(y=insurance$expenses , p=0.70, list=FALSE)
train_input <- insurance[inTrain,-7]
test_input <- insurance[-inTrain,-7]
# Create an improved explanatory model using interaction terms on the train data set 
ins_improved_train_model1 <- lm(train_target ~ age + children + bmi + sex + bmi30*smoker + region, data = train_input)
summary(ins_improved_train_model1)

```

# Create an improved explanatory model using both interaction and quadratic terms on the train data set 

```{r Train an improved lm model}
# add the quadratic term
ins_improved_train_model2 <- lm(train_target ~ age + age2+ children + bmi + sex + smoker+ region, data = train_input)
summary(ins_improved_train_model2)

# add both interaction and quadratic terms
ins_improved_train_model3 <- lm(train_target ~ age + age2 + children + bmi + sex + bmi30*smoker + region, data = train_input)
summary(ins_improved_train_model3)
```

# Generate prediction and evaluation metrics from the improved trained lm model  

```{r Apply and evaluate an improved lm model}
# generate prediction on the testing data with the improved trained model 
predictions_improved_test <- predict(ins_improved_train_model3, test_input)
mmetric(test_target,predictions_improved_test,c("MAE","RMSE","MAPE","RMSPE","RAE", "RRSE","COR", "R2"))
# generate prediction on the training data with the improved trained model 
predictions_improved_train <- predict(ins_improved_train_model3, train_input)
mmetric(train_target,predictions_improved_train,c("MAE","RMSE","MAPE","RMSPE","RAE", "RRSE","COR", "R2"))
```
# cross validation (cv) for numeric predictions

a. setup cv parameters and create folds
b. 3-fold cv of lm using lapply and anonymous function
c. display cv results using kable; generate and display means and sds of metrics
d. define a named function - cv_function
e. run cv_function for 3-fold and 10-fold cross validatons of predictive models

```{r set up cv parameters and create folds}

# Use the original insurance data
# Remove age2 and bmi30
insurance$age2 <- NULL
insurance$bmi30 <- NULL

#  Set up cv parameters
# df identifies the whole data set by its name
# target identifies the target variable by its column index in df
# nFolds indicates the number of folds for cv
# seedVal carries the seed value for random sampling of instances when creating folds
# prediction_method indicates the prediction method - e.g., lm
# metric_list is a list of evaluation metrics that mmetric should generate

df <- insurance
target <- 7
nFolds <- 3
seedVal <- 500
prediction_method <- lm
# This is the same as above: assign("prediction_method", lm)
metrics_list <- c("MAE","RMSE","MAPE","RMSPE","RAE","RRSE","R2")

# create folds using the assigned values

set.seed(seedVal)
folds = createFolds(df[,target],nFolds)
str(folds)
# elements in folds: folds$Fold1 = folds[[1]], folds$Foldi = folds[[i]]
# length of folds = nFolds
```

# cross validation using lapply and anonymous function

```{r cross validation using lapply and anonymous function}

# Input to lapply - folds, output returned from lapply - cv_results have the same list length - i.e., nFolds
# Input to function(x) is x=folds[[i]]
# For each possible value of x, execute the list of commands in {} after function(x)

cv_results <- lapply(folds, function(x)
{ 
# data preparation:
# vertically (i.e., row-wise) divide df using x (a fold) and -x  into test and train
# horizontally (i.e., column-wise) divide df using target and -target into target and input columns

  test_target <- df[x,target]
  test_input <- df[x,-target]
  
  train_target <- df[-x,target]
  train_input <- df[-x,-target]
   
  pred_model <- prediction_method(train_target~.,train_input)
  # compare the trained linear regression model by fold
  print(summary(pred_model))
  
  pred <- predict(pred_model, test_input)
# return saves performance results in cv_results[[i]]
  return(mmetric(test_target,pred,metrics_list))
})

cv_results
```

# Display cv performances results in a table using kable() in knitr package

```{r show cv_results, means and sds using kable}

# convert a list to a data frame using as.data.frame and convert this data frame to a matrix before using rowSds()
str(as.matrix(cv_results))
str(as.matrix(as.data.frame(cv_results)))
cv_results_m <- as.matrix(as.data.frame(cv_results))

# Generate and show Means and Sds of performance metrics over all folds
# Need matrixStats package
# save rowMeans and rowSds results in matrices which are the required input format for the use of colnames()

cv_mean<- as.matrix(rowMeans(cv_results_m))
cv_sd <- as.matrix(rowSds(cv_results_m))

# ?rowSds, ?colnames and cbind to learn more about rowSds, colnames and cbind
colnames(cv_mean) <- "Mean"
colnames(cv_sd) <- "Sd"

# Combine and show cv_results and Means and Sds
cv_all <- cbind(cv_results_m, cv_mean, cv_sd)

# t(cv_all) transposes cv_all
kable(t(cv_all),digits=2)
```

# cross validation using a named function

```{r define cv_function}

cv_function <- function(df, target, nFolds, seedVal, prediction_method, metrics_list)
{
  # create folds
  set.seed(seedVal)
  folds = createFolds(df[,target],nFolds) 
  # perform cross validation
  cv_results <- lapply(folds, function(x)
  { 
    test_target <- df[x,target]
    test_input  <- df[x,-target]

    train_target <- df[-x,target]
    train_input <- df[-x,-target]

    prediction_model <- prediction_method(train_target~.,train_input) 
    pred<- predict(prediction_model,test_input)
    return(mmetric(test_target,pred,metrics_list))
  })
  # generate means and sds and show cv results, means and sds using kable
  cv_results_m <- as.matrix(as.data.frame(cv_results))
  cv_mean<- as.matrix(rowMeans(cv_results_m))
  cv_sd <- as.matrix(rowSds(cv_results_m))
  colnames(cv_mean) <- "Mean"
  colnames(cv_sd) <- "Sd"
  cv_all <- cbind(cv_results_m, cv_mean, cv_sd)
  kable(t(cv_all),digits=2)
}
```

# 3-fold and 10 fold cross validations of lm using cv_function

```{r run cv_function with lm}

# df <- insurance
# target <- 7
# nFolds <- 3
# seedVal <- 500
# metrics_list <- c("MAE","RMSE","MAPE","RMSPE","RAE","RRSE","R2")
# assign("prediction_method", lm)

### call cv_function for 3-fold and 10-fold cvs of lm
cv_function(df, target, 3, seedVal, prediction_method, metrics_list)
cv_function(df, target, 10, seedVal, prediction_method, metrics_list)
# The metric means are similar between 3-fold and 10-fold cvs. 
# Hence, 3-fold cv is sufficient to estimate the means of performance metrics.
# The sds of metrics increase from 3-fold to 10-fold cv.
# Hence, increase nFolds in cv help assess the grow of sds with nFolds.

```

# 3-fold and 10 fold cross validations of rpart using cv_function

```{r run cv_function with rpart}

# assign("prediction_method", rpart)

cv_function(df, target, 3, seedVal, rpart, metrics_list)
cv_function(df, target, 10, seedVal, rpart, metrics_list)
# Due to non-linear relationships, rpart models improve over the baseline lm models on the basis of each metric
# The means of rpart models' metrics are slightly different between 10-fold and 3-fold cvs. Higher nFolds might be necessary for cv of rpart models.
```

# 3-fold and 10 fold cross validations of M5P using cv_function

```{r run cv_function with M5P}

# assign("prediction_method", M5P)

cv_function(df, target, 3, seedVal, M5P, metrics_list)
cv_function(df, target, 10, seedVal, M5P, metrics_list)
# M5P models further improve over rpart models. The performance is similar to lm models improved with quadratic and interaction terms
### end of tutorial
```

